---
title: "Class11_Part 4_from_Class_10"
format: pdf
author: Dhruv 
---

## **Part 4 (11/6/24)**

Comparative Structural Analysis 

```{r}
library(bio3d)
```

```{r}
id <- "1ake_A"
```

```{r}
aa <- get.seq(id)
aa
```
```{r}
b <- blast.pdb(aa)
```
```{r}
head(b$hit.tbl)
```
```{r}
hits <- plot(b)
```

What is inside "hits"

```{r}
attributes(hits)
```
What are the top hits form our blast results 

```{r}
hits$pdb.id
```

Download PDB files
```{r}
files <- get.pdb(hits$pdb.id, path = "pdbs", split = TRUE, gzip = TRUE)
```

I have now found and downloaded all ADK structures in the PDB database. Viewing them is difficult and needs to be alligned and then superimposed. 
This is evident when we use MolViewer to load upp all the structures mentioned before. 

Install BiocManager package from CRAN. use `BiocManager::install()` to install any package from bioconductor. 

Align releated PDBs

```{r}
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
pdbs
```

Now that we have our sequences aligned, we can now do a PCA

```{r}
pc <- pca(pdbs)
plot(pc, pc.axes = c(1:2))
```

To examine what the PC is capturing, we can plot the loadings or make a movie (trajector) of moving along PC1

```{r}
mktrj(pc, pc=1, file="pc1.pdb")
```



